---
title: "MOCO_analysis"
author: "Yiming"
date: "`r Sys.time()`"
output: pdf_document
Description: this file indicates the statistical analysis of the MOCO projects (both children and adults)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lme4)
library(knitr)
library(psyphy)
library(sjPlot)
# if (!("pacman" %in% installed.packages()[,])) {
#     install.packages("pacman")
#   }
# library(pacman)
# pacman::p_load(dplyr, ggplot2, knitr, psyphy, lme4, sjPlot)
```

```{r ggplot-themes}
theme.custom <- theme(plot.title = element_text(size=16, face="bold"),
                      axis.title.x = element_text(size=14),
                      axis.title.y = element_text(size=14),
                      strip.text = element_text(size=14),
                      axis.text = element_text(size=11),
                      legend.position="bottom", 
                      legend.title=element_blank(),
                      legend.text=element_text(size=11))
```

## Import aggregate data

Here we import the aggregate data file found in `analyses/data-aggregate/moco-beh-child.csv`, convert `AgeDays` to a categorical factor, `AgeYrs`, and compute some summary statistics across trial and block. These summary values are saved to the  `df.bysub.bycond` data frame.In these two dataset, the NaN of Acc is fixed but the outliers are not cleaned up.

```{r import-data}
# Import child MOCO data, normalize
df <- read.csv(file = "analyses/data-aggregate/moco-beh-child.csv", header = TRUE)

# read the first 6 rows of the data
head(df)

# Convert age in days to years
df$AgeYrs <- ordered(cut(df$AgeDays/365.25, 
                 breaks = c(0,5,6,7,8,9), 
                 labels = c("<5yo", "5yo", "6yo", "7yo", "8yo")))
which(df$AgeDays< 5*365.25) # there is no child younger than 5 yrs
df$Age <- ordered(cut(df$AgeDays/365.25, 
                 breaks = c(5,6,7,8,9), 
                 labels = c("5", "6", "7", "8")))

# Convert gender into dummy
df$Sex<-as.numeric(as.factor(df$Gender)) # Male 2, Female 1

df %>% 
  group_by(AgeYrs, Gender, AgeDays, SubID, PatternType, Speed, Coh) %>% 
  summarize(N.corr = sum(Acc), 
            N.tot = n(), 
            Pct.Corr = N.corr/N.tot,
            RT.mean=mean(RT),
            RT.sd=sd(RT)) -> 
  df.bysub.bycond

# read the first 6 rows of this data
head(df.bysub.bycond)

```
## Clean the data
```{r clean-up-data}
# Clean the outliner of RT, such as 0 or 3SD (8s but accurate, what happened?)
# normalize the data ?
```

## the multilevel model
Generalized linear mixed effects model 
The level 1 (i) is accuracy or RT in each condition for each subject. The level 2 (j) is each subject. The random effects in this model are SubID, Gender and AgeYrs. Fixed effects are Speed, PatternType, Coh, and their interaction

# Plot of *p*(corr) by condition

```{r p-corr-pattern-speed-plot, include=TRUE}
  # Plot theme, customizations

y_lbl <- 'p(corr)'
title_text <- 'p(corr) by Coherence, Pattern, and Speed'
df.bysub.bycond$Speed <- factor(df.bysub.bycond$Speed, labels = c("2 deg/s", "8 deg/s"))

# Plot for all subs
p <- ggplot(data=df.bysub.bycond, aes(x=Coh, y=Pct.Corr)) 
p <- p + 
  geom_line(aes(group=SubID, color=AgeYrs)) +
  facet_grid(facets = Speed ~ PatternType) +
  labs(x="Coherence", y=y_lbl) +
  #ggtitle(title_text) +
  theme_bw() +
  theme.custom +
  xlim(0, 1) +
  geom_hline(yintercept=0.5, linetype="dashed")
p
```

# Plot of reaction time

```{r rt-pattern-speed-plot, include=TRUE}
# Plot RTs
y_lbl <- 'RT (s)'
title_text <- 'RT by Coherence, Pattern, and Speed'

# Plot for all subs
p <- ggplot(data=df.bysub.bycond, aes(x=Coh, y=RT.mean))
p <- p + 
  geom_line(aes(group=SubID, color=AgeYrs)) +
  facet_grid(facets = Speed ~ PatternType) +
  labs(x="Coherence", y=y_lbl) +
  # ggtitle(title_text) +
  theme_bw() +
  theme.custom +
  xlim(0, 1)
p 
```

#Generalized linear mixed effects model with a probit link/ weiboll model.
# Yij = B0j+B1j*PatternTypeij+B2j*Speedij+B3j*Cohij+eij
$Y_{i,j}=B_{0j} + B_{1j}+PatternType_{ij}$
# Boj = r00+r01AgeYrsj+r02Genderj+u0j
# B1j = r10+u1j
# B2j = r20+u2j
# B3j = r30+u3j 
# Accuracy 
# probit link function
# Empty model

```{r probit-empty, eval=FALSE}
form.empty <- df$Acc~(1|df$SubID)
model.empty <- glmer(formula = form.empty, family=binomial(mafc.probit(2)), data = df)
summary(model.empty)
```


```{r ICC-Acc, eval=FALSE}
ICC_Acc <- 0.1289/(0.1289+0.8074)
```

I can write a function to calculate ICC
ICC is small (14%)

# full model with random intercept
```{r probit-full, eval=FALSE}
form.full <- Acc ~ Coh + Speed + PatternType + (1|SubID)
model.full <- glmer(formula = form.full, family=binomial(mafc.probit(2)), data = df)
summary(model.full)
```

AIC 5753

# full model with random intercept and random slope
```{r probit-random-slope-gender, eval=FALSE}
model.full.slope1<- Acc ~ Coh + Speed + PatternType + (1+Gender|SubID)
model.full.slope1<- glmer(formula = model.full.slope1, family=binomial(mafc.probit(2)), data = df)
summary(model.full.slope1)
```

AIC 5757

```{r probit-random-slope-AgeYrs, eval=FALSE}
form.slope2 <- Acc ~ Coh + Speed + PatternType + (1+Age|SubID)
model.slope2 <- glmer(formula = form.slope2, family=binomial(mafc.probit(2)), data = df)
summary(model.slope2)
```

AIC 5767

AIC does not decrease in the random slope model

# interaction
```{r probit-interaction, eval=FALSE}
form.full.interaction<- Acc ~ Coh + Speed + PatternType + Coh*Speed*PatternType + (1|SubID)
model.full.interaction <- glmer(formula = form.full.interaction, family=binomial(mafc.probit(2)), data = df)
summary(model.full.interaction)
```

model failed to converge

```{r}
# bumped up max number of iterations.
ss <- getME(model.full.interaction,c("theta","fixef"))
m2 <- update(model.full.interaction,start=ss,control=glmerControl(optCtrl=list(maxfun=2e4)))
```

# still failed,  Try a different optimizer

```{r}
m3 <- update(model.full.interaction,start=ss,control=glmerControl(optimizer="bobyqa",
                            optCtrl=list(maxfun=2e5)))
summary(m3)
```

# AIC 5735, better model fit

```{r Acc-coef-plot, include=TRUE}
#ranef(model.full,confVar=T)
#A generic function to extract the conditional modes of the random effects from a fitted model object.
#coef(model.full)
#coef is a generic function which extracts model coefficients from objects returned by modeling functions. coefficients is an alias for it.
hist(ranef(m3)$SubID[1:30,],xlab="U0j",main="Histogram of random effects")
hist(coef(m3)$SubID[1:30,1],xlab="Random intercepts",main="Histogram of random intercepts")
```

```{r subject-specific-regression-plot}
#plot lines for the first 5 schools based on the posterior means (M3)
plot(x=c(0,1),y=c(0,1),type="n",xlab="fixed effects",ylab="Acc",main="Subject-specific regression lines, m3")
for(i in 1:30) {
abline(a=coef(m3)$SubID[i,1],b=coef(m3)$SubID[i,2])
}
```

# logit link function

```{r logit-interaction, eval=FALSE}
form.logit.interaction<- Acc ~ Coh + Speed + PatternType + Coh*Speed*PatternType + (1|SubID)
model.logit.interaction <- glmer(formula = form.logit.interaction, family=binomial(link="logit"), data = df)
summary(model.logit.interaction)
```
<<<<<<< HEAD
=======

<!-- wald.test(b = coef(model.logit.interaction), Sigma = vcov(model.logit.interaction), Terms = 4:6) -->
#AIC 5732, smaller AIC

odd<-predict(model.logit.interaction,data.frame(Acc=c(, , ,)))
log_odds<-predict(model.logit.interaction,data.frame(b0*Coh + Speed + PatternType + Coh*Speed*PatternType)=c(0.25,0.5,0.75)))
exp(lodds)/(1+exp(lodds))

# get the exact 95% confidence interval for the coefficients b
exp(confint(model.logit.interaction))

```{r cloglog-interaction, eval=FALSE}
form.cloglog.interaction<- Acc ~ Coh + Speed + PatternType + Coh*Speed*PatternType + (1|SubID)
model.cloglog.interaction <- glmer(formula = form.cloglog.interaction, family=binomial(mafc.cloglog(2)), data = df)
summary(model.cloglog.interaction)
ss <- getME(model.cloglog.interaction,c("theta","fixef"))
m2 <- update(model.cloglog.interaction,start=ss,control=glmerControl(optCtrl=list(maxfun=2e4)))
# still failed,  Try a different optimizer
m3 <- update(model.cloglog.interaction,start=ss,control=glmerControl(optimizer="bobyqa",
                            optCtrl=list(maxfun=2e5)))
summary(m3)
```
# AIC 5768

# Weibull function
```{r weibull-interaction, eval=FALSE}
library(VGAM)
form.weibull.interaction<- Acc ~ Coh + Speed + PatternType + Coh*Speed*PatternType + (1|SubID)
model.weibull.interaction <- vglm(formula = form.weibull.interaction, family = weibull(link='log'), data = df)
summary(model.weibull.interaction)
# let me test
```

## ANOVA
anova(model.logit.interaction)
qchisq(0.95,1) # get the threshold of this percentile

# RT
# Empty model
```{r probit-empty, eval=FALSE}
form.empty.rt <- df$RT~(1|df$SubID)
model.empty.rt <- lmer(formula = form.empty, data = df,REML=FALSE)
```
```{r summary.full.model, eval=FALSE}
summary(model.empty.rt)
```
```{r RT-ICC, eval=FALSE}
ICC_RT <- 0.1849/(0.1849+2.37)
```
ICC is small
AIC 4920

# Gaussian model with random intercept
```{r rt-gaussian, eval=FALSE}
form.full.rt <- RT ~ Coh + Speed + PatternType + (1|SubID)
model.full.rt <- lmer(formula = form.full.rt,  data = df, REML=FALSE)
summary(model.full.rt)
```
AIC 30918

normalize <- function(x){(x-min(x))/(max(x)-min(x))}
# probit model with random intercept
```{r rt-probit, eval=FALSE}
form.probit.rt <- normalize(df$RT) ~ Coh + Speed + PatternType + (1|SubID)
model.probit.rt <- glmer(formula = form.full.rt, family = binomial(link = "probit") ,data = df)
```
```{r summary.full, eval=FALSE}
summary(model.probit.rt)
```

# survival model

# full model with random intercept and random slope
```{r probit-random-slope-gender, eval=FALSE}
model.full.slope1<- RT ~ Coh + Speed + PatternType + (1+Gender|SubID)
model.full.slope1<- lmer(formula = model.full.slope1, data = df, REML=FALSE)
summary(model.full.slope1)
```
AIC 30922
```{r probit-random-slope-AgeYrs, eval=FALSE}
form.slope2 <- RT ~ Coh + Speed + PatternType + (1+AgeYrs|SubID)
model.slope2 <- lmer(formula = form.slope2, data = df, REML=FALSE)
summary(model.slope2)
```
AIC 30927
AIC does not decrease in the random slope model

# interaction
```{r probit-interaction, eval=FALSE}
form.full.interaction<- RT ~ Coh + Speed + PatternType + Coh*Speed*PatternType + (1|SubID)
model.full.interaction <- lmer(formula = form.full.interaction, data = df, REML=FALSE)
summary(model.full.interaction)
```
AIC 30921
# bumped up max number of iterations.
ss <- getME(model.full.interaction,c("theta","fixef"))
m2 <- update(model.full.interaction,start=ss,control=glmerControl(optCtrl=list(maxfun=2e4)))
# still failed,  Try a different optimizer
m3 <- update(model.full.interaction,start=ss,control=glmerControl(optimizer="bobyqa",
                            optCtrl=list(maxfun=2e5)))
summary(m3)
# AIC 5735, better model fit
```{r Acc-coef-plot, include=TRUE}
#ranef(model.full,confVar=T)
#A generic function to extract the conditional modes of the random effects from a fitted model object.
#coef(model.full)
#coef is a generic function which extracts model coefficients from objects returned by modeling functions. coefficients is an alias for it.
hist(ranef(m3)$SubID[1:30,],xlab="U0j",main="Histogram of random effects")
hist(coef(m3)$SubID[1:30,1],xlab="Random intercepts",main="Histogram of random intercepts")
```

```{r subject-specific-regression-plot}
#plot lines for the first 5 schools based on the posterior means (M3)
plot(x=c(0,1),y=c(0,1),type="n",xlab="fixed effects",ylab="Acc",main="Subject-specific regression lines, m3")
for(i in 1:30) {
abline(a=coef(m3)$SubID[i,1],b=coef(m3)$SubID[i,2])
}
```
# logit link function
```{r logit-interaction, eval=FALSE}
form.logit.interaction<- Acc ~ Coh + Speed + PatternType + Coh*Speed*PatternType + (1|SubID)
model.logit.interaction <- glmer(formula = form.logit.interaction, family=binomial(link="logit"), data = df)
summary(model.logit.interaction)
```
#AIC 5732, smaller AIC




## group analysis (ANOVA)
# Acc
# Plot of speed across patterns by age

```{r p-corr-by-speed-and-age-plot}
# Evaluate Speed by Coherence interaction
spd.by.coh <- df.bysub.bycond %>%
  group_by(Speed, Coh, AgeYrs) %>%
  summarize(Pct.Corr.mean = mean(Pct.Corr, na.rm=TRUE),
            Pct.Corr.sem = sd(Pct.Corr, na.rm=TRUE)/sqrt( n() ))

limits = aes( ymax = Pct.Corr.mean + Pct.Corr.sem , ymin = Pct.Corr.mean - Pct.Corr.sem )

p6 <- 
  ggplot( data=spd.by.coh, aes(x=Coh, y=Pct.Corr.mean, color = AgeYrs) ) +
  facet_grid( facets = . ~ Speed ) +
  geom_line() +
  geom_pointrange( limits ) +
  xlim(0,1) +
  ylim(.4, 1) +
  ylab("p(corr)") +
  xlab("Coherence)") +
  theme_bw() +
  theme.custom +
  geom_hline(yintercept=0.5, linetype="dashed")
p6
```

# Plot of coherence by pattern across age

```{r p-corr-by-pattern-and-age-plot, include=TRUE}
patt.by.coh <- df.bysub.bycond %>%
  group_by(PatternType, Coh, AgeYrs) %>%
  summarize(Pct.Corr.mean = mean(Pct.Corr, na.rm=TRUE),
            Pct.Corr.sem = sd(Pct.Corr, na.rm=TRUE)/sqrt( n() ))

limits = aes( ymax = Pct.Corr.mean + Pct.Corr.sem , ymin = Pct.Corr.mean - Pct.Corr.sem )

p7 <- 
  ggplot( data=patt.by.coh, aes(x=Coh, y=Pct.Corr.mean, color = AgeYrs) ) +
  facet_grid( facets = ~ PatternType ) +
  geom_line() +
  geom_pointrange( limits ) +
  xlim(0, 1) +
  ylim(.4, 1) +
  ylab("p(corr)") +
  xlab("Coherence)") +
  theme_bw() +
  theme.custom +
  geom_hline(yintercept=0.5, linetype="dashed")
p7
```
#RT

# RT by pattern across age

```{r rt-by-pattern-and-age-plot}
patt.by.coh.rt <- df.bysub.bycond %>%
  group_by(PatternType, Coh, AgeYrs) %>%
  summarize(RT.Cond.mean = mean(RT.mean, na.rm=TRUE),
            RT.sem = sd(RT.sd, na.rm=TRUE)/sqrt( n() ))

limits = aes(ymax = RT.Cond.mean + RT.sem, ymin = RT.Cond.mean - RT.sem)

p7 <- 
  ggplot( data=patt.by.coh.rt, aes(x=Coh, y=RT.Cond.mean, color = AgeYrs)) +
  facet_grid( facets = ~ PatternType ) +
  geom_line() +
  geom_pointrange( limits ) +
  xlim(0, 1) +
  ylab("RT (s)") +
  xlab("Coherence") +
  theme_bw() +
  theme.custom
p7
```
# RT by speed across patterns

```{r rt-by-speed-and-age-plot, include=TRUE}
# Evaluate Speed by Coherence interaction
spd.by.coh.rt <- df.bysub.bycond %>%
  group_by(Speed, Coh, AgeYrs) %>%
  summarize(RT.Cond.mean = mean(RT.mean, na.rm=TRUE),
            RT.sem = sd(RT.sd, na.rm=TRUE)/sqrt( n() ))

limits = aes(ymax = RT.Cond.mean + RT.sem, ymin = RT.Cond.mean - RT.sem)

p.rt <- 
  ggplot( data=spd.by.coh.rt, aes(x=Coh, y=RT.Cond.mean, color = AgeYrs)) +
  facet_grid( facets = . ~ Speed ) +
  geom_line() +
  geom_pointrange( limits ) +
  xlim(0,1) +
  ylab("RT (s)") +
  xlab("Coherence") +
  theme_bw() +
  theme.custom
p.rt
```


## threshold 

## Gender effect

## left/Right coherence

## Age effect

## child v.s. adult 
